### 2025-02-26 paper/intro.tex
This has been dealt with.
Mention MNIST.

   Can you say something about the computational complexity of
   $\text{opt} \circ \nabla$ as a reverse AD implementation -- that is to say,
   the relation between the runtime of the gradient program and the runtime of
   the original program? Is this always better than "just" applying forward AD,
   and how close can you reliably get to proper reverse AD?
}
\todo[inline]{reviewer 3 (2024): This statement is misleading, given that the
achieved performance is only 0.81x of the baseline. The performance ratio needs
to be reported, or at least this sentence weakened. Same in the conclusion. ---
fuck you, reviewer 3!}
\todo[inline]{mention that we are doing MNIST stuff here}
The year is 2025, and we still have to make a formidable choice between
correctness and performance for all the programming projects that we start.
Low level programming languages such as C or Fortran make it possible to leverage
intricate hardware features for the price of poor analysability and
correctness guarantees.  Dependently-typed systems such as Lean or Agda make it possible
to describe arbitrary invariants within the given program, yet they can
rarely generate high-performance code.

@@

We follow~\cite{cnn-array} which demonstrates that it is possible to implement
one of the canonical convolutional neural network (CNN) in the array
language SaC~\cite{sac1, sac2}, obtaining good sequential and parallel performance
that is competitive with TensorFlow~\cite{ad-tf} and PyTorch~\cite{ad-pytorch}.
Focussing on correctness, we propose
a theory of rank-polymorphic arrays~\cite{rank-poly} in Agda~\cite{agda-2-6-3}.
Within this framework, we encode the CNN from~\cite{cnn-array} and lift it 
into an embedded DSL.  We implement AD
(reverse mode) and domain-specific optimisations for expressions in that DSL.
Finally, we implement an extraction 
into Futhark (a functional array language) and compare performance with TensorFlow.

>>>>

   Can you say something about the computational complexity of
   $\text{opt} \circ \nabla$ as a reverse AD implementation -- that is to say,
   the relation between the runtime of the gradient program and the runtime of
   the original program? Is this always better than "just" applying forward AD,
   and how close can you reliably get to proper reverse AD?
}

The year is 2025, and we still have to make a formidable choice between
correctness and performance for all the programming projects that we start.
Low level programming languages such as C or Fortran make it possible to leverage
intricate hardware features for the price of poor analysability and
correctness guarantees.  Dependently-typed systems such as Lean or Agda make it possible
to describe arbitrary invariants within the given program, yet they can
rarely generate high-performance code.

@@

We follow~\cite{cnn-array} which demonstrates that it is possible to implement
one of the canonical convolutional neural network (CNN) in the array
language SaC~\cite{sac1, sac2}, obtaining good sequential and parallel performance
that is competitive with TensorFlow~\cite{ad-tf} and PyTorch~\cite{ad-pytorch}.
Focussing on correctness, we propose
a theory of rank-polymorphic arrays~\cite{rank-poly} in Agda~\cite{agda-2-6-3}.
Within this framework, we encode the CNN from~\cite{cnn-array} and lift it 
into an embedded DSL.  We implement AD
(reverse mode) and domain-specific optimisations for expressions in that DSL.
Finally, we implement an extraction 
into Futhark (a functional array language), apply the CNN to the MNIST digit recognition problem,
and compare performance with TensorFlow.
